---
title: "R Notebook"
output: html_notebook
---
The purpose of this project is to compare forecasting methodologies on data sets in Introductory Time Series with R

We want to be able to do both:

* Decompose series into trend/seasonality(??)/other
* Forecast k-periods ahead and evaluate pred error and confidence interval


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

```{r}
library(dplyr)
library(ggplot2)
library(tseries)
library(forecast)
path_data = "http://www.maths.adelaide.edu.au/andrew.metcalfe/Data/"


# plot(maine_data)

```

```{r}

file_data = "Maine.dat"
file_path = file_data %>% paste0(path_data,.)
maine_data = read.table(file_path,header = TRUE)



file_data = "cbe.dat"
file_path = file_data %>% paste0(path_data,.)
cbe_data = read.table(file_path,header = TRUE)


file_data = "pounds_nz.dat"
file_path = file_data %>% paste0(path_data,.)
GBPNZD = read.table(file_path,header = TRUE)


file_data = "global.dat"
file_path = file_data %>% paste0(path_data,.)
global_temp = read.table(file_path,header = TRUE)

file_data = "wave.dat"
file_path = file_data %>% paste0(path_data,.)
wave = read.table(file_path,header = TRUE)

file_data = "Fontdsdt.dat"
file_path = file_data %>% paste0(path_data,.)
Font = read.table(file_path,header = TRUE)

file_data = "wine.dat"
file_path = file_data %>% paste0(path_data,.)
wine = read.table(file_path,header = TRUE)

file_data = "HP.txt"
file_path = file_data %>% paste0(path_data,.)
HP = read.table(file_path,header = TRUE)

file_data = "stemp.dat"
file_path = file_data %>% paste0(path_data,.)
stemp = read.table(file_path,header = TRUE)


AP = data("AirPassengers")

maine_ts  = ts(maine_data$unemploy,start=c(1996,1),freq=12)
maine_ts
```

```{r}
list_data = c()


AP
maine_ts
global_temp
GBPNZD
maine_data
cbe_data
wave
Font
wine
HP
stemp
```

```{r}
maine_ts %>% plot
```

```{r}
# maine_annual = maine_ts %>% aggregate()/12
# maine_annual

dfr_data = maine_data %>%
  mutate(timestampp = 1:n() ) 

dfr_data %>% ggplot2::ggplot(ggplot2::aes(x=timestampp, y=unemploy)) +
ggplot2::geom_point() +
ggplot2::geom_smooth()


```

```{r}

#model
ts_len = dfr_data %>% nrow()
train_frac = 0.7

num_train = (ts_len * train_frac) %>% floor()

num_test = ts_len - num_train
num_test

```
```{r}


train_idx = dfr_data[num_train, 'timestampp']
print(train_idx)
dfr_data_new =dfr_data %>% 
  mutate(train_test = ifelse(timestampp < train_idx, 'train','test'))
```

```{r}
# ------------  Holt-Winters
mod_HW$fitted[,1] %>% as.numeric()
dfr_data_new %>% head


# -------- TRAINING ---------------------------------

#fit HoltWinters - this will fit trend and seasonal
mod_HW =
  dfr_data_new %>%
  dplyr::filter(train_test == 'train') %>%
    select(unemploy) %>%
    ts(start = c(1996,1),freq=12) %>%
    HoltWinters()

#Add HW prediction error back on

#get offset to HW data - there is 'burn in' like a moving average with HW model so have to offset the data
number_training = mod_HW$x %>% nrow
number_HW_fitted = mod_HW$fitted %>% nrow
offset_fitted_data = number_training - number_HW_fitted 

#create df with fitted data, correct ts, and labels
dfr_data_pred = dfr_data_new[(offset_fitted_data+1):(train_idx-1),] %>% 
  mutate(unemploy = mod_HW$fitted[,1] %>% as.numeric()) %>%
  mutate(train_test = 'HW_predict')

#adding HW fitted data back to data
dfr_data_with_preds = dfr_data_new %>% 
  bind_rows(dfr_data_pred) 

#plot both training data and HW fitted data
dfr_data_with_preds %>%ggplot(aes(x=timestampp, y=unemploy, colour= train_test)) +
  geom_line()

####Error analysis on training data, can we improve model by forecasting error

#calc prediction error
pred_error =   dfr_data_with_preds %>% 
  dplyr::filter((train_test == "train") | (train_test == "HW_predict")) %>%
  spread(train_test,unemploy) %>%
  mutate(error = train - HW_predict)

#plot error
pred_error %>% ggplot(aes(x=timestampp, y = error)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0)

#get error vector
error = pred_error%>% 
  drop_na() %>% 
  # select(error) %>%
  dplyr::pull(error)

#fit ARIMA to error
model_error_auto = error %>% auto.arima
model_error_auto
model_error = error %>% arima(order= c(1,0,1))
model_error$coef
model_error_auto$fitted


# -------- FORECASTING ---------------------------------

#we want to make n-step ahead forecasts

#we do forecast of determenistic level, trend, season using n-step ahead HW
#we can use ARIMA model to forecast error paths and use these to get confidence intervals
#then check does pred fall in confidence interval

pred = mod_HW %>% 
  predict(n.ahead = num_test ) %>% 
  as.numeric()

# dfr_data_frame = 
dfr_data_pred =  dfr_data_new[(train_idx):(ts_len-1),] %>%
 mutate(unemploy = pred) %>%
 mutate(train_test = 'HW_predict')

dfr_data_with_preds = dfr_data_new %>% 
  bind_rows(dfr_data_pred) 

dfr_data_with_preds %>%ggplot(aes(x=timestampp, y=unemploy, colour= train_test)) +
  geom_line()


```

```{r}
library(tidyr)
pred_error =   dfr_data_with_preds %>% 
  dplyr::filter((train_test == "test") | (train_test == "HW_predict")) %>%
  spread(train_test,unemploy) %>%
  mutate(error = test - HW_predict)
  

pred_error %>% ggplot(aes(x=timestampp, y = error)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0)


#check integration - looks pretty integrated
#plot acf

#Actually check prediction error looks close to train error

error = pred_error%>% 
  drop_na() %>% 
  # select(error) %>%
  dplyr::pull(error)

error %>% adf.test


error %>% acf #doesn't show integrated

#fit an error model to error in training data




error %>% pacf

library(forecast)
model_error_auto = error %>% auto.arima
model_error = error %>% arima(order= c(1,0,1))
model_error$coef
model_error_auto$fitted


# adf.test(pred_error%>% drop_na())

```

