---
title: "R Notebook"
output: html_notebook
---
The purpose of this project is to compare forecasting methodologies on data sets in Introductory Time Series with R

We want to be able to do both:

* Decompose series into trend/seasonality(??)/other
* Forecast k-periods ahead and evaluate pred error and confidence interval


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

```{r}
library(dplyr)
library(ggplot2)
library(tseries)
library(forecast)
library(tidyr)

path_data = "http://www.maths.adelaide.edu.au/andrew.metcalfe/Data/"


# plot(maine_data)

```

```{r}

file_data = "Maine.dat"
file_path = file_data %>% paste0(path_data,.)
maine_data = read.table(file_path,header = TRUE)



file_data = "cbe.dat"
file_path = file_data %>% paste0(path_data,.)
cbe_data = read.table(file_path,header = TRUE)


file_data = "pounds_nz.dat"
file_path = file_data %>% paste0(path_data,.)
GBPNZD = read.table(file_path,header = TRUE)


file_data = "global.dat"
file_path = file_data %>% paste0(path_data,.)
global_temp = read.table(file_path,header = TRUE)

file_data = "wave.dat"
file_path = file_data %>% paste0(path_data,.)
wave = read.table(file_path,header = TRUE)

file_data = "Fontdsdt.dat"
file_path = file_data %>% paste0(path_data,.)
Font = read.table(file_path,header = TRUE)

file_data = "wine.dat"
file_path = file_data %>% paste0(path_data,.)
wine = read.table(file_path,header = TRUE)

file_data = "HP.txt"
file_path = file_data %>% paste0(path_data,.)
HP = read.table(file_path,header = TRUE)

file_data = "stemp.dat"
file_path = file_data %>% paste0(path_data,.)
stemp = read.table(file_path,header = TRUE)


AP = data("AirPassengers")

maine_ts  = ts(maine_data$unemploy,start=c(1996,1),freq=12)
maine_ts
```

```{r}
list_data = c()


AP
maine_ts
global_temp
GBPNZD
maine_data
cbe_data
wave
Font
wine
HP
stemp
```

```{r}
maine_ts %>% plot
```

```{r}
# maine_annual = maine_ts %>% aggregate()/12
# maine_annual

dfr_data = maine_data %>%
  mutate(timestampp = 1:n() ) 

dfr_data %>% ggplot2::ggplot(ggplot2::aes(x=timestampp, y=unemploy)) +
ggplot2::geom_point() +
ggplot2::geom_smooth()


```

```{r}

#model
ts_len = dfr_data %>% nrow()
train_frac = 0.8

num_train = (ts_len * train_frac) %>% floor()

num_test = ts_len - num_train
num_test

```
```{r}


train_idx = dfr_data[num_train, 'timestampp']
print(train_idx)
dfr_data_new =dfr_data %>% 
  mutate(train_test = ifelse(timestampp < train_idx, 'train','test'))
```

```{r}
# ------------  Holt-Winters
mod_HW$fitted[,1] %>% as.numeric()
dfr_data_new %>% head


# -------- TRAINING ---------------------------------

#fit HoltWinters - this will fit trend and seasonal
mod_HW =
  dfr_data_new %>%
  dplyr::filter(train_test == 'train') %>%
    select(unemploy) %>%
    ts(start = c(1996,1),freq=12) %>%
    HoltWinters()

#Add HW prediction error back on

#get offset to HW data - there is 'burn in' like a moving average with HW model so have to offset the data
number_training = mod_HW$x %>% nrow
number_HW_fitted = mod_HW$fitted %>% nrow
offset_fitted_data = number_training - number_HW_fitted 

#create df with fitted data, correct ts, and labels
dfr_data_pred = dfr_data_new[(offset_fitted_data+1):(train_idx-1),] %>% 
  mutate(unemploy = mod_HW$fitted[,1] %>% as.numeric()) %>%
  mutate(train_test = 'HW_predict')


#adding HW fitted data back to data
dfr_data_with_preds = dfr_data_new %>% 
  bind_rows(dfr_data_pred) 

#plot both training data and HW fitted data
dfr_data_with_preds %>%ggplot(aes(x=timestampp, y=unemploy, colour= train_test)) +
  geom_line()

####Error analysis on training data, can we improve model by forecasting error

#calc prediction error
pred_error =   dfr_data_with_preds %>% 
  dplyr::filter((train_test == "train") | (train_test == "HW_predict")) %>%
  spread(train_test,unemploy) %>%
  mutate(error = train - HW_predict)

#plot error
pred_error %>% ggplot(aes(x=timestampp, y = error)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0)
#we want to make n-step ahead forecasts

#we do forecast of determenistic level, trend, season using n-step ahead HW
#we can use ARIMA model to forecast error paths and use these to get confidence intervals
#then check does pred fall in confidence interval

pred = mod_HW %>% 
  predict(n.ahead = num_test ) %>% 
  as.numeric()

# dfr_data_frame = 
dfr_data_pred =  dfr_data_new[(train_idx):(ts_len-1),] %>%
 mutate(unemploy = pred) %>%
 mutate(train_test = 'HW_predict')

dfr_data_with_preds = dfr_data_new %>% 
  bind_rows(dfr_data_pred) 

dfr_data_with_preds %>%ggplot(aes(x=timestampp, y=unemploy, colour= train_test)) +
  geom_line()

mod_HW$fitted %>% plot
```

```{r}
#Training ARIMA
model_ARIMA_auto = dfr_data_new %>%
  dplyr::filter(train_test == 'train') %>%
    select(unemploy) %>%
    ts(start = c(1996,1),freq=12) %>%
    auto.arima()
model_ARIMA_auto



#Forecasting ARIMA

pred = model_ARIMA_auto %>% 
  predict(n.ahead = num_test ) %$% #weird operator shorthand for with(.)
  pred %>%
  as.numeric()

# dfr_data_frame = 
dfr_data_pred =  dfr_data_new[(train_idx):(ts_len-1),] %>%
 mutate(unemploy = pred) %>%
 mutate(train_test = 'ARIMA_predict')

dfr_data_with_preds = dfr_data_new %>% 
  bind_rows(dfr_data_pred) 

dfr_data_with_preds %>%ggplot(aes(x=timestampp, y=unemploy, colour= train_test)) +
  geom_line()

```

```{r}
fit_model = function(model_function, dfr_in, pc_train){
  
  #model
  ts_len = dfr_data %>% nrow()
  train_frac = 0.8
  
  num_train = (ts_len * train_frac) %>% floor()
  
  num_test = ts_len - num_train
  num_test
  
  train_idx = dfr_data[num_train, 'timestampp']
  
  dfr_data_new =dfr_data %>% 
    mutate(train_test = ifelse(timestampp < train_idx, 'train','test'))
  
  # ------------  Holt-Winters
  mod_HW$fitted[,1] %>% as.numeric()
  dfr_data_new %>% head
  
  
  # -------- TRAINING ---------------------------------
  
  ##################
  # define model_fit_function(df_train, target_col, prediction_label ,parameters) -> list(model_inner_fitted, dfr_out)
  ##################
  #
  #fit HoltWinters - this will fit trend and seasonal
  mod =
    dfr_data_new %>%
    dplyr::filter(train_test == 'train') %>%
      select(unemploy) %>%
      ts(start = c(1996,1),freq=12) %>%
      HoltWinters()
  
  #Add HW prediction error back on
  
  #get offset to HW data - there is 'burn in' like a moving average with HW model so have to offset the data
  number_training = mod_HW$x %>% nrow
  number_HW_fitted = mod_HW$fitted %>% nrow
  offset_fitted_data = number_training - number_HW_fitted 
  
  #create df with fitted data, correct ts, and labels
  dfr_data_pred = dfr_data_new[(offset_fitted_data+1):(train_idx-1),] %>% 
    mutate(unemploy = mod_HW$fitted[,1] %>% as.numeric()) %>%
    mutate(train_test = 'HW_predict')
  
  
  #adding HW fitted data back to data
  dfr_data_with_preds = dfr_data_new %>% 
    bind_rows(dfr_data_pred) 
  
  ##################
  #define model_predict(model, dfr_in, prediction_label, steps_ahead) -> dfr_out
  ##################
  
    #we do forecast of determenistic level, trend, season using n-step ahead HW
    #we can use ARIMA model to forecast error paths and use these to get confidence intervals
    #then check does pred fall in confidence interval
    
    pred = mod_HW %>% 
      predict(n.ahead = num_test ) %>% 
      as.numeric()
    
    # dfr_data_frame = 
    dfr_data_pred =  dfr_data_new[(train_idx):(ts_len-1),] %>%
     mutate(unemploy = pred) %>%
     mutate(train_test = 'HW_predict')
    
    dfr_data_with_preds = dfr_data_new %>% 
      bind_rows(dfr_data_pred) 
  
  output_res = list(model, dfr_out)
  
  return(model_results)
}
```

Note:
Forecasts from both
* Holt-Winters
* ARIMA(seasonal) 
Seem to be really sensitive to trend in data
The forecasts seem to predict a trend that follows the last few periods of the data.
E.g: 
* If training has upwards direction in trend component near end of data, then forecasts follow that
* HW seems less susceptible on a couple of trials

Reasons:
* HW fitting routine in R is designed to mimise the one-step ahead errors, maybe we need to minimise multiple steps ahead
* auto ARIMA fits in sample data and doesn't worry about forecasting at all in optimisation precedure, seems a bit wrong


TODO:
* put these in a function
* try different data set lengths
* test over different data sets and write insights
* do state space model
* we are interested in forecasting multiple steps ahead, so want to pick models that minimise that error, not the in sample training error
  * need to work out k-fold cross val equivalent for timeseries
  * grid search the parameters for ARIMA (urgh)
  * optimise for 


```{r}

### Holt-Winters Bootstrapping training data

# we can simulate from the error data model to create N more data sets, then re-fit the HW model, to get estimated C.I and mean of parameters

#we can the forecast ahead with each model,use the bootstrapped parameter estimates + simulated ARIMA errors

#then we see how our test data looks vs training

#get error vector
error = pred_error%>% 
  drop_na() %>% 
  select(error) #%>%
  # as.numeric()

#fit ARIMA to error
model_error_auto = error %>% auto.arima
model_error_auto
model_error = error %>% arima(order= c(1,0,1))
model_error$coef
model_error_auto$fitted



```

```{r}
# -------- FORECASTING ---------------------------------





```

```{r}
pred_error =   dfr_data_with_preds %>% 
  dplyr::filter((train_test == "test") | (train_test == "HW_predict")) %>%
  spread(train_test,unemploy) %>%
  mutate(error = test - HW_predict)
  

pred_error %>% ggplot(aes(x=timestampp, y = error)) +
  geom_point() +
  geom_line() +
  geom_hline(yintercept = 0)


#check integration - looks pretty integrated
#plot acf

#Actually check prediction error looks close to train error

error = pred_error%>% 
  drop_na() %>% 
  # select(error) %>%
  dplyr::pull(error)

error %>% adf.test


error %>% acf #doesn't show integrated

#fit an error model to error in training data




error %>% pacf

library(forecast)
model_error_auto = error %>% auto.arima
model_error = error %>% arima(order= c(1,0,1))
model_error$coef
model_error_auto$fitted


# adf.test(pred_error%>% drop_na())

```

